```{r setup,include=FALSE}
# set the knitr options ... for everyone!
# if you unset this, then vignette build bonks. oh, joy.
#opts_knit$set(progress=TRUE)
opts_knit$set(eval.after='fig.cap')
# for a package vignette, you do want to echo.
# opts_chunk$set(echo=FALSE,warning=FALSE,message=FALSE)
opts_chunk$set(warning=FALSE,message=FALSE)
#opts_chunk$set(results="asis")
opts_chunk$set(cache=TRUE,cache.path="cache/")

#opts_chunk$set(fig.path="github_extra/figure/",dev=c("pdf","cairo_ps"))
#opts_chunk$set(fig.path="github_extra/figure/",dev=c("png","pdf"))
opts_chunk$set(fig.path="github_extra/figure/",dev=c("png"))
opts_chunk$set(fig.width=5,fig.height=4,dpi=64)

# doing this means that png files are made of figures;
# the savings is small, and it looks like shit:
#opts_chunk$set(fig.path="figure/",dev=c("png","pdf","cairo_ps"))
#opts_chunk$set(fig.width=4,fig.height=4)
# for figures? this is sweave-specific?
#opts_knit$set(eps=TRUE)

# this would be for figures:
#opts_chunk$set(out.width='.8\\textwidth')
# for text wrapping:
options(width=64,digits=2)
opts_chunk$set(size="small")
opts_chunk$set(tidy=TRUE,tidy.opts=list(width.cutoff=50,keep.blank.line=TRUE))

#PDQutils.meta <- packageDescription('PDQutils')
```

# PDQutils

Approximating Probability Distributions.

-- Steven E. Pav, shabbychef@gmail.com

## Installation

This package may be installed from CRAN; the latest version may be
found on [github](https://www.github.com/shabbychef/PDQutils "PDQutils")
via devtools:

```{r install,eval=FALSE,echo=TRUE}
if (require(devtools)) {
	# latest greatest
	install_github('shabbychef/PDQutils')
}
```

# Basic Usage

Approximating the distribution of a random variable via the Gram Charlier or Cornish Fisher
expansions is most convenient when the random variable can be decomposed as the sum of a 
small number of independent random variables whose cumulants can be computed. For example, 
suppose $Y = \sum_{1 \le i \le k} \sqrt{X_i / \nu_i}$ where the $X_i$ are independent central 
chi-square random variables with degrees of freedom $\nu_1,\nu_2,...,\nu_k$. I will call this
a 'snak' distribution, for 'sum of Nakagami', since each summand follows a 
[Nakagami distribution](https://en.wikipedia.org/wiki/Nakagami_distribution "Nakagami distribution").
We can easily write code that generates variates from this distribution given a vector
of the degrees of freedom:

```{r rsnak,eval=TRUE,echo=TRUE}
rsnak <- function(n,dfs) {
	samples <- Reduce('+',lapply(dfs,function(k) { sqrt(rchisq(n,df=k)/k) }))
}
```

Let's take one hundred thousand draws from this distribution and see whether it is approximately normal:

```{r testit,eval=TRUE,echo=TRUE,dpi=200,out.width='600px',out.height='500px'}
n.samp <- 1e5
dfs <- c(8,15,4000,10000)
set.seed(18181)
# now draw from the distribution 
rvs <- rsnak(n.samp,dfs)
qqnorm(rvs)
qqline(rvs,col='red')
```

Using the additivity
property of cumulants, we can compute the cumulants of $Y$ easily if we have the cumulants of
the $X_i$. These in turn can be computed from the raw moments.  See
[wikipedia](https://en.wikipedia.org/wiki/Chi_distribution "chi distribution") for the raw moments
of the Chi distribution. The following function computes the cumulants:

```{r snakcu,eval=TRUE,echo=TRUE}
# for the moment2cumulant function:
library(PDQutils)
# compute the first ord.max raw cumulants of the sum of Nakagami variates
snak_cumulants <- function(dfs,ord.max=10) {
	# first compute the raw moments
	moms <- lapply(dfs,function(k) { 
		ords <- 1:ord.max
		moms <- 2 ^ (ords/2.0) * exp(lgamma((k+ords)/2) - lgamma(k/2))
		# we are dividing the chi by sqrt the d.f.
		moms <- moms / (k ^ (ords/2.0))
		moms
	})
	# turn moments into cumulants
	cumuls <- lapply(moms,moment2cumulant)
	# sum the cumulants
	tot.cumul <- Reduce('+',cumuls)
	return(tot.cumul)
}
```

We can now implement the 'dpq' functions trivially using the Gram-Charlier and Cornish-Fisher
approximations, as follows:

```{r dpqsnak,eval=TRUE,echo=TRUE}
library(PDQutils)

dsnak <- function(x,dfs,ord.max=10,...) {
	raw.moment <- cumulant2moment(snak_cumulants(dfs,ord.max))
	retval <- dapx_gca(x,raw.moment,...)
	return(retval)
}
psnak <- function(q,dfs,ord.max=10,...) {
	raw.moment <- cumulant2moment(snak_cumulants(dfs,ord.max))
	retval <- papx_gca(q,raw.moment,...)
	return(retval)
}
qsnak <- function(p,dfs,ord.max=10,...) {
	raw.cumul <- snak_cumulants(dfs,ord.max)
	retval <- qapx_cf(p,raw.cumul,...)
	return(retval)
}
```

The qqplot should look better now:

```{r improvedqq,eval=TRUE,echo=TRUE,dpi=200,out.width='600px',out.height='500px'}

qqplot(qsnak(ppoints(n.samp),dfs=dfs),rvs,
	main='Q-Q against qsnak (C-F appx.)')
qqline(rvs, distribution=function(p) qsnak(p,dfs),
	col='red')

```


